---
title: "Datasets Available in R"
author: "Saghir Bashir"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    highlight: default
    code_folding: show
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
> *Note:* To search for R datasets please download (or clone) this github repository and open the *WhatData.html* file in your browser.

## Motivation & Objective

Often my main objective is to find an R datasets of a particular class (e.g. `tbl_df`, `data.frame`, `ts`) which I would like to use for: 

* teaching
* presentations 
* to prepare reproducible example code for sharing (see `reprex` package)

There *are* many useful datasets available in Base R and others that are loaded with packages. To see a list of those that are available to you (including for loaded packages) type `data()` in the R console. There are packages dedicated to sharing data (e.g. see `gapminder`, `FiveThirtyEight` packages) which are not covered in this document.

This program creates a `tibble` call `all_ds` that makes it easier to search for datasets based on the keywords and/or classes. 

## A tibble of R Datasets

Start by loading the required packages and those from which you would like to include datasets. 
```{r echo=TRUE, message=FALSE}
library(tidyverse)
library(tibble)
library(stringr)
library(MASS)
library(DT)
```

Creating a `tibble` (`all_ds`) with the list of datasets and other useful information. It is intentionally in _long_ format as it makes it easier to search by class. You can use `dplyr::filter` to search `all_ds` or alternatively use the `DT::datatable` below.
```{r echo=TRUE}
all_ds_ls <- data()

all_ds <- all_ds_ls$results %>%
          tibble::as_tibble() %>%
          dplyr::rename(Data_Orig = Item) %>%
          dplyr::mutate(dataset = stringr::word(Data_Orig, 1)) %>%
          dplyr::mutate(Class = purrr::invoke_map(class, as.list(parse(text = dataset)))) %>%
          tidyr::unnest(Class) %>%
          dplyr::select(Package, Data_Orig, Title, Class) %>%
          dplyr::arrange(Package, Data_Orig, Class)
```

## Find R Datasets

### Interactive Data Table

Simply type in a key word, class or package name into the data table "*Search:*" box (see below)
```{r echo=TRUE}
# Quick and lazy method ;o)
DT::datatable(all_ds)
```

### Using dplyr::filter 

```{r eval=FALSE}
all_ds %>% 
  filter(str_detect(Class, "tbl_df"))

all_ds %>% 
  filter(str_detect(Title, regex("sleep", ignore_case = TRUE)))

all_ds %>% 
  filter(str_detect(Title, regex("econ|stock|share", ignore_case = TRUE)) &
                  str_detect(Class, "ts"))
```

**Warning:** This document is intentionally in HTML format otherwise the (`DT::datatable`) search will not work.