---
title: "Datasets Available in R"
author: "Saghir Bashir"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    highlight: default
    code_folding: show
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(DT)
```
> *Note:* To search for R datasets please download (or clone) this github repository and open the *WhatData.html* file in your browser.

## Motivation & Objective

There *are* many useful datasets available in Base R and the packages that are installed on 
your system. Sometimes you might need to find an R datasets of a particular class  (e.g.
`tbl_df`, `data.frame`, `ts`) that you can use for: 

* teaching,
* a presentation or
* to prepare reproducible example code for sharing (see `reprex` package).

Based on the packages that are installed [WhatData](https://github.com/saghirb/WhatData) creates 
a `data.frame` (called `all_ds`) of datasets which can be searched using keywords and/or classes. 

## A data.frame (tibble) of R Datasets

Creating a dataset with the list of available datasets and other useful information. 
```{r allData, echo=TRUE, message=FALSE, warning=FALSE}
# Function to catch the error for data that is not exported (although it is
# returned using the 'data()' function).
unexportedData <- function (x) {
  out <- tryCatch(class(eval(parse(text = x))), error = function(e) "NOT EXPORTED")
  return(out)
}

all_ds <- data(package = .packages(all.available = TRUE)) %>% 
  .$results %>%
  tibble::as_tibble() %>%
  dplyr::mutate(DataOrig = stringr::word(Item, 1)) %>%
  dplyr::mutate(pkgData = paste(Package, DataOrig, sep="::")) %>% 
  dplyr::arrange(pkgData) %>% 
  dplyr::mutate(Class = purrr::invoke_map(unexportedData, pkgData)) %>% 
  tidyr::unnest(Class) %>% 
  dplyr::filter(!str_detect(Class, "NOT EXPORTED")) %>% 
  dplyr::select(pkgData, Package, DataOrig, Title, Class) %>%
  dplyr::arrange(pkgData, Class) %>%
  dplyr::mutate(Val = Class) %>%
  tidyr::spread(key = Class, value=Val, fill = "") %>%
  tidyr::unite(Classes, c(-pkgData, -Package, -DataOrig, -Title), sep= " ")
```

## Find R Datasets

### Interactive Data Table

Simply type in a key word, class or package name into the data table "*Search:*" box (see below)
```{r dataDT, echo=TRUE}
all_ds %>% 
  select(-pkgData) %>% 
  DT::datatable(filter = "top")
```

### Using dplyr::filter 

```{r how2Serach, eval=FALSE}
# Find all tibbles.
all_ds %>% 
  filter(str_detect(Classes, "tbl_df"))

# Filtering out the rows with class "ts" and not things like "datasets".
all_ds %>% 
  filter(str_detect(Classes, regex("\\b(ts)\\b")))

# Find datasets related with any form of "sleep" in dataset names or descriptions.
all_ds %>% 
  filter(str_detect(Title, regex("sleep", ignore_case = TRUE)))

# Time series data for econ(omic), stock or share data.
all_ds %>% 
  filter(str_detect(Title, regex("econ|stock|share", ignore_case = TRUE)) &
                  str_detect(Classes, regex("\\b(ts)\\b")))
```

**Note:** This document is intentionally in HTML format otherwise the (`DT::datatable`) search will not work.