---
title: "Datasets Available in R"
author: "Saghir Bashir"
date: "`r Sys.Date()`"
output: html_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What Data?

There are many datasets avalable in R that can be used for teaching, presentations or to prepare reproducible example code for sharing (see `reprex` package). To see a list of the available datasets (including for the loaded packages) type `data()` in the R console. 

This program creates a dataset that makes it easier to search for dataset to use based on the dataset descriptions and classes.

## Dataset of R Datasets

First load the packages for which we would like to include in our dataset of R datasets. 
```{r echo=FALSE}
library(tidyverse)
library(tibble)
library(stringr)
library(DT)
```

Creating a `tibble` with the list of datasets and other useful information.
```{r echo=FALSE}
all_ds_ls <- data()

all_ds <- all_ds_ls$results %>%
          as_tibble() %>%
          rename(Data_Orig = Item) %>%
          mutate(dataset = word(Data_Orig, 1)) %>%
          select(Package, Data_Orig, dataset, Title) %>%
          mutate(Class = invoke_map(class, as.list(parse(text = dataset)))) %>%
          unnest(Class)
```

## Making the dataset accessible

```{r echo=FALSE}

whatDat <- function(search=""){
  
}

# Quick and lazy method ;o)
datatable(all_ds)
```

